---
title: "Green Carpet Distribution"
author: "Paul J Palmer"
date: "21/11/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load some libraries that are needed but not yet transferred into the package
library(ggmap)
library(tidyverse)
library(broom)
library(ggpolypath)
library(GreenCarpetProjectFiles)

```

```{r include=FALSE}
library(sp)
library(raster)
library(broom)
library(ggpolypath)
# This code makes a plogon stencil to mask the edge of the map.
bb <- unlist(attr(leics,"bb"))
# Note in the online example I based this on coords were transposed!
coords <- cbind(
  bb[c(1,3,3,1)],
  bb[c(2,2,4,4)]
)

sp <- SpatialPolygons(
  list(Polygons(list(Polygon(coords)), "id")), 
  proj4string = CRS(proj4string(leics)))
sp_diff <- erase(sp, leics)
sp_diff_df <- fortify(sp_diff) # This is a stencil with a leicestershire shape hole.

```

## R Setup Code


First load the data

```{r Get-data, echo=TRUE, message=FALSE, warning=FALSE }
# Note that we are running this implicitly from the data-raw directory
#Green.Carpet.Data <- readRDS("Green.Carpet.Data.Rds")
data("Green.Carpet.Data")
data("leics")
```



## A first plot

```{r A-first-plot, echo=FALSE, include=TRUE}
#this is a test
data("leics")
Phenology.Plot("Green Carpet All Data", Green.Carpet.Data$DOY)
data("leics")
```


Note the data has three peaks showing multiple flight times for the the imago.

```{r A-first-map, echo=FALSE, include=TRUE}
Plot.Maps("All data", Green.Carpet.Data$Gridref, leics)
```


The map shows the taxon distributed generally across the whole county. However, this type of presence-absence map does not show the density of the distribution.

## An improved representation including density

The following map uses mulitple layers including a probability densty distribution. This does not reflect underlying variances in recording effort, but it does give a better sense of the variation in distribution.

```{r echo=FALSE, include=TRUE, warning=FALSE,message=FALSE,error=FALSE}
# lat.long <- GBNG.to.Lat.Long(Green.Carpet.Data$Gridref)
 data.to.plot <- Green.Carpet.Data
leics_df <- tidy(leics) # Need to load broom
#ggmap(leics_df)
#ggplot() +
 #   geom_path(data = leics_df, aes(x = long, y = lat, group = group)) +
#    labs(title = "ggplot map of roads")



 my.map <- ggplot() + geom_path(data = leics_df, aes(x = long, y = lat, group = group)) + 
  stat_density_2d(  aes( x = Long, y = Lat, fill = stat(level)  ), size = 2, bins = 10, data = data.to.plot, geom = "polygon", contour = TRUE, show.legend = FALSE) +
  scale_fill_gradient(low = "grey", high= "black") +
    geom_point(data = data.to.plot, aes(x = Long, y = Lat)) + geom_polygon(data = leics_df, fill=NA, colour="black", aes(x = long, y = lat, group = group)) +
   theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) +
    geom_polypath(
     aes( long, lat, group=group),
     sp_diff_df,
     fill="white",
     alpha=1
   )  
 
 
 my.map
#Plot.Maps.ggplot("Green Carpet Post 2010 Sep", data.to.plot$Gridref, leics)
 # Create a box covering the whole by retrieving the bounding box for leics
 
 
```


## Changes in distribution through time

### Pre 1980

Pre 1980 the taxon was noted to be single brooded and tended to be confined to the Northern part of the county.

```{r Pre-1980, echo=FALSE, include=TRUE}
pre.1980 <- Green.Carpet.Data[Green.Carpet.Data$YYYY < 1980,]

# Now drop the rows for which there are no Dates
rows.with.dates <- grepl("\\d{4}-\\d{2}-\\d{2}",pre.1980$YYMMDD )
rows.to.keep <- which(rows.with.dates)
pre.1980 <-pre.1980[rows.to.keep,]
Phenology.Plot("Green Carpet Pre 1980", pre.1980$DOY)

```


```{r echo=FALSE, include=TRUE, warning=FALSE,message=FALSE,error=FALSE}
lat.long <- GBNG.to.Lat.Long(pre.1980$Gridref)
data.to.plot <-pre.1980
leics_df <- tidy(leics) # Need to load broom
#ggmap(leics_df)
#ggplot() +
 #   geom_path(data = leics_df, aes(x = long, y = lat, group = group)) +
#    labs(title = "ggplot map of roads")



 my.map <- ggplot() + geom_path(data = leics_df, aes(x = long, y = lat, group = group)) + 
  stat_density_2d(  aes( x = Long, y = Lat, fill = stat(level)  ), size = 2, bins = 5, data = data.to.plot, geom = "polygon", contour = TRUE, show.legend = FALSE) +
  scale_fill_gradient(low = "grey", high= "black") +
    geom_point(data = data.to.plot, aes(x = Long, y = Lat)) + geom_polygon(data = leics_df, fill=NA, colour="black", aes(x = long, y = lat, group = group)) +
   theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) +
    geom_polypath(
     aes( long, lat, group=group),
     sp_diff_df,
     fill="white",
     alpha=1
   )  
 
 
 my.map
#Plot.Maps.ggplot("Green Carpet Post 2010 Sep", data.to.plot$Gridref, leics)
 # Create a box covering the whole by retrieving the bounding box for leics
 
 
```

## Post 1980

Post 1980 the picture has become more complex. The taxon is double brooded with a smaller September occurance. Note that the distribution is more Southerly.


```{r echo=FALSE, include=TRUE}

Post.1980 <- Green.Carpet.Data[Green.Carpet.Data$YYYY > 1980,]

# Now drop the rows for which there are no Dates
rows.with.dates <- grepl("\\d{4}-\\d{2}-\\d{2}",Post.1980$YYMMDD )
rows.to.keep <- which(rows.with.dates)
Post.1980 <-Post.1980[rows.to.keep,]

Phenology.Plot("Green Carpet Post 1980", Post.1980$DOY)

```

```{r echo=FALSE, include=TRUE, warning=FALSE,message=FALSE,error=FALSE}
#lat.long <- GBNG.to.Lat.Long(Post.1980$Gridref)
data.to.plot <- Post.1980
leics_df <- tidy(leics) # Need to load broom
#ggmap(leics_df)
#ggplot() +
 #   geom_path(data = leics_df, aes(x = long, y = lat, group = group)) +
#    labs(title = "ggplot map of roads")



 my.map <- ggplot() + geom_path(data = leics_df, aes(x = long, y = lat, group = group)) + 
  stat_density_2d(  aes( x = Long, y = Lat, fill = stat(level)  ), size = 2, bins = 6, data = data.to.plot, geom = "polygon", contour = TRUE, show.legend = FALSE) +
  scale_fill_gradient(low = "grey", high= "black") +
    geom_point(data = data.to.plot, aes(x = Long, y = Lat)) + geom_polygon(data = leics_df, fill=NA, colour="black", aes(x = long, y = lat, group = group)) +
   theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) +
    geom_polypath(
     aes( long, lat, group=group),
     sp_diff_df,
     fill="white",
     alpha=1
   )  
 
 
 my.map
#Plot.Maps.ggplot("Green Carpet Post 2010 Sep", data.to.plot$Gridref, leics)
 # Create a box covering the whole by retrieving the bounding box for leics
 
 
```

The chart above shows all of the post 1980 records

```{r echo=FALSE, include=TRUE, warning=FALSE,message=FALSE,error=FALSE}



# Now drop the rows for which there are no Dates
#rows.with.dates <- grepl("\\d{4}-\\d{2}-\\d{2}",Green.Carpet.Data$YYMMDD )
#rows.to.keep <- which(rows.with.dates)
#Post.1980.Dates <- Green.Carpet.Data[rows.to.keep,]

Post.1980.Dates <- Post.1980[strtoi(Post.1980$MM) == 7,]
Post.1980.Dates <- na.omit(Post.1980.Dates)

#lat.long <- GBNG.to.Lat.Long(Post.1980.Dates$Gridref)
data.to.plot <-Post.1980.Dates
leics_df <- tidy(leics) # Need to load broom
#ggmap(leics_df)
#ggplot() +
 #   geom_path(data = leics_df, aes(x = long, y = lat, group = group)) +
#    labs(title = "ggplot map of roads")



 my.map <- ggplot() + geom_path(data = leics_df, aes(x = long, y = lat, group = group)) + 
  stat_density_2d(  aes( x = Long, y = Lat, fill = stat(level)  ), size = 2, bins = 5, data = data.to.plot, geom = "polygon", contour = TRUE, show.legend = FALSE) +
  scale_fill_gradient(low = "grey", high= "black") +
    geom_point(data = data.to.plot, aes(x = Long, y = Lat)) + geom_polygon(data = leics_df, fill=NA, colour="black", aes(x = long, y = lat, group = group)) +
   theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) +
    geom_polypath(
     aes( long, lat, group=group),
     sp_diff_df,
     fill="white",
     alpha=1
   )  
 
 
 my.map

```

The chart above shows the June records post 1980 which represent the single brooded population  The more records occur in Rutland than pre 1980.

```{r echo=FALSE, include=TRUE, warning=FALSE,message=FALSE,error=FALSE}



# Now drop the rows for which there are no Dates
#rows.with.dates <- grepl("\\d{4}-\\d{2}-\\d{2}",Green.Carpet.Data$YYMMDD )
#rows.to.keep <- which(rows.with.dates)
#Post.1980.Dates <- Green.Carpet.Data[rows.to.keep,]

Post.1980.Dates <- Post.1980[strtoi(Post.1980$MM) == 5 ,]
Post.1980.Dates <- na.omit(Post.1980.Dates)

#lat.long <- GBNG.to.Lat.Long(Post.1980.Dates$Gridref)
data.to.plot <- Post.1980.Dates
leics_df <- tidy(leics) # Need to load broom
#ggmap(leics_df)
#ggplot() +
 #   geom_path(data = leics_df, aes(x = long, y = lat, group = group)) +
#    labs(title = "ggplot map of roads")



 my.map <- ggplot() + geom_path(data = leics_df, aes(x = long, y = lat, group = group)) + 
  stat_density_2d(  aes( x = Long, y = Lat, fill = stat(level)  ), size = 2, bins = 5, data = data.to.plot, geom = "polygon", contour = TRUE, show.legend = FALSE) +
  scale_fill_gradient(low = "grey", high= "black") +
    geom_point(data = data.to.plot, aes(x = Long, y = Lat)) + geom_polygon(data = leics_df, fill=NA, colour="black", aes(x = long, y = lat, group = group)) +
   theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) +
    geom_polypath(
     aes( long, lat, group=group),
     sp_diff_df,
     fill="white",
     alpha=1
   )  
 
 
 my.map

```

In May the taxon is more frequently seen in the south west.
